All Samples Model

### 2018-03-12 AllSamplesModel.Rmd
...

source('lib/load_patient_metadata.R')
source('lib/cv-pt-glm.R')
source('lib/common_plots.R')

@@

cache.dir = paste(data, 'Analysis/5e6_arms_all', sep='/')

@@

@@

ggplot(patient.info, aes(Maximal,Circumference, color=Status)) + geom_point() + labs(title='C vs M') + plot.theme

@@

gp = globalCN.plot + theme(legend.position = 'none') + labs(y='mean adj. seg.', title='All samples')
ggsave(filename='plots/globalCN.png', plot=gp, width=16, height=4.6, units='in', dpi=300)
gp

gbe = globalBECN.plot + theme(legend.position = 'none') + labs(y='mean adj.seg.', title="Non-Dysplastic Barrett's Samples") 
ggsave(filename='plots/globalBECN.png', plot=gbe, width=16, height=4.6, units='in', dpi=300)
gbe

rm(locs, gp, gbe)
```

@@

tt = t.test(subset(x, Status == 'Progressor')$cx, subset(x, Status == 'Non-Progressor')$cx)
pv = ggplot(x, aes(Status, cx, fill=Status)) + 
  geom_jitter(width=0.2, color='grey39') + 
  geom_boxplot(outlier.colour = NA, alpha=0.8) +
  labs(x='', y='cx', title='Per sample cx scores', subtitle=paste('p-value=', format.pval(tt$p.value, digits=2))) + theme(legend.position = 'none') +
  plot.theme
pv

@@

```{r noHGD, echo=F, message=F, warning=F, fig.height=16, fig.width=16, eval=T}
rm(performance.at.1se, coefs, mp)

file = paste(cache.dir, 'nohgd.Rdata', sep='/')
if (file.exists(file)) {
  message(paste("loading file", file))
  load(file, verbose=T)
} else {
  stop(paste("No model file", file, "found"))
}
nohgd.coefs = coefs[[select.alpha]]
#do.call(grid.arrange, c(no.hgd.plots, top="No HGD/IMC samples", ncol=2))
ggsave('plots/nohgd_eyebrow.png', plot=do.call(grid.arrange, c(no.hgd.plots, top="No HGD/IMC samples", ncol=2)), height=10, width=10, units='in', dpi=300)
```

```{r, echo=F, fig.width=6, fig.height=6}
mp = model.performance(performance.at.1se, coefs)
mp$plot = mp$plot + labs(title='No HGD/IMC samples')
mp$plot

@@

```{r noLGD, echo=F, warning=F, message=F, fig.height=16, fig.width=16}
file = paste(cache.dir, 'nolgd.Rdata', sep='/')
if (file.exists(file)) {
  load(file, verbose=T)
} else {
  stop(paste("No model file", file, "found"))
}

#do.call(grid.arrange, c(nolgd.plots, top="No LGD/HGD/IMC samples", ncol=2))
```

```{r, echo=F, fig.width=6, fig.height=6}
mp = model.performance(performance.at.1se, coefs)
mp$plot = mp$plot + labs(title='No LGD/HGD/IMC samples')
mp$plot

@@

```{r loonohgd, echo=F, message=F, warning=T}
file = paste(cache.dir, 'loonohgd.Rdata', sep='/')
if (file.exists(file)) {
  message(paste("loading file", file))
  load(file, verbose=T)
} else {
  stop(paste("No file", file, "found"))
}
```

```{r, echo=F, message=F, fig.width=5, fig.height=5}
pg.sampNOHGD = do.call(rbind, pg.sampNOHGD)

@@

```{r, echo=F, fig.width=6, fig.height=6}
ggplot(pg.sampNOHGD, aes(Prediction)) + geom_histogram(aes(fill=..x..), breaks=seq(0,1,0.1), show.legend = F) + scale_fill_distiller(palette = 'RdYlBu', name='P(P)') + labs(title='Prediction using model w/o HGD', y='n Samples', x='Probability') + plot.theme

@@

#ggsave('plots/haz_coef.png', plot=p, height=5, width=5, units='in', dpi=300)

@@

```{r nohgd_coefs, message=F, warning=F, echo=F}
setdiff(ch$label, rownames(nohgd.coefs))

@@

#ggsave('plots/global_cn_coefs.png', plot=p, width=16, height=4.6, units='in', dpi=300)

@@

#ggsave('plots/ndbe_cn_coefs.png', plot=p, width=16, height=4.6, units='in', dpi=300)

@@

oddPt = names(pg.samp)[which.max(performance.at.1se)]
subset(sum.patient.data, Hospital.Research.ID == oddPt)
pg.samp[[oddPt]]$Prediction

oddPt = names(pg.samp)[which.min(performance.at.1se)]

@@

load(file=paste(cache.dir, 'insilico_splits.Rdata', sep='/'), verbose=T)

@@

@@

@@

@@

#grid.arrange( tableGrob(annotation.table, rows=NULL, theme=tt3) )
pred.confidence$Risk = factor(pred.confidence$Risk, levels=c('Low','Moderate','High'), ordered = T)
```

```{r pcalib, echo=F, warning=F, message=F, fig.width=8, fig.height=6}
pcal = ggplot(pred.confidence, aes(mn, perc)) + 
  geom_rect(aes(xmin=r1, xmax=r2, ymin=0,ymax=1, fill=Risk), alpha=0.6) + scale_fill_manual(values=riskCols) +
  geom_point() + geom_abline(slope=1, intercept=0, color='grey39', linetype='dotted') + #geom_text(aes(label=paste(r1,r2,sep='-')), nudge_x=0.05) +
  geom_errorbar(aes(ymin=ci.low, ymax=ci.high), width=0.01) + labs(x='Mean prediction', y='Ratio of `P` patients', title='Prediction calibration') +
  scale_color_manual(values=riskCols) + plot.theme + theme(legend.position = 'none')
pcal

@@

mod = subset(back, Risk == 'Moderate')

@@

unadjRR = ggplot(back, aes(OR)) + geom_histogram(aes(fill=..x..), bins=15, show.legend = F) +
  #scale_fill_distiller(palette = 'RdYlBu', type='div',name='P(P)') +
  scale_fill_gradientn(colors = myPal,  name='') + 
  labs(y='n Samples', x='Relative Risk', title='Unadjusted relative risk') + plot.theme
unadjRR

ggsave(filename='plots/unadj_relrisk.png', plot=unadjRR, width=6, height=5, units='in', dpi=300)

@@

ggsave(filename='plots/feature_jacknife.png', plot=p, width=5, height=6, unit='in', dpi=300)

@@

rp = ggplot(riskPath, aes(Pathology, ratio*100)) + geom_bar(aes(group=Risk, fill=Risk),stat='identity', position=position_stack()) + 
  scale_fill_manual(values=riskCols) + scale_color_manual(values=c('white','black','white')) +
  geom_text(data=pathTotal, aes(label=paste('n=',nPath,sep=''), x=Pathology, y=101.5), position=position_stack(), size=4) +
  geom_text(data=subset(riskPath, ratio>0), aes(group=Risk,label=paste(ratio*100,'%',sep=''), color=Risk), position=position_stack(vjust = 0.5), size=5, show.legend = F) +
  facet_grid(~Status, scales='free_x', space = 'free_x') + plot.theme + labs(title='Samples predicted by pathology', y='% Predicted', x='Pathology') + theme(legend.position = 'bottom')
rp

@@

# ggplot(all.preds, aes(Patient, value/n.samples, group=variable, fill=variable)) + 
#   geom_bar(stat='identity', position = position_stack()) +
#   scale_fill_manual(values=riskCols) + 
#   geom_text(aes(y=1.01, label=n.samples)) +
#   facet_wrap(~Status, scales='free_x') + labs(y='Sample ratios', title='Per patient ratio of predictions') + 
#   plot.theme

@@

endos = back %>% group_by(Status, months.before.final, brks, PID) %>% dplyr::summarise(
  'pred.endo'=ifelse(length(which(Risk == 'High')) > 0, 1, 0), 
  'n.endos'=length(unique(PID)), 
  'n.samples'=length(PID),
  'mean.pred'=mean(Prediction),
  'pred.samples'=length(which(Risk == 'High')),
  'max.path'=max(Pathology))

endos$quant =  cut(endos$mean.pred, breaks = cuts)
endos = merge(endos, pred.confidence[c('quant', 'Risk')])

sq = c(0, 1, 1*12, 2*12, 4*12, 6*12, 8*12, 10*12, 15*12)
endos$brks = cut(endos$months.before.final, include.lowest = T, ordered_result = T,
breaks=sq,
labels=c('Endpoint','1-12','12-24','24-48','48-72','72-96','96-120','>120'))
endos$brks = with(endos, factor(brks, rev(levels(brks)))) # invert order

endos.m = subset(endos, Status == 'P') %>% group_by(brks) %>% dplyr::summarise( 
  'pred.r'=sum(pred.endo)/sum(n.endos),'n.samp'=sum(n.samples), 'n.endo'=sum(n.endos))

tmp = subset(back, Status == 'P')
tmp$brks = cut(tmp$months.before.final, include.lowest = T, ordered_result = T,breaks=sq,
labels=c('Endpoint','1-12','12-24','24-48','48-72','72-96','96-120','>120'))
tmp = tmp %>% group_by(brks) %>% dplyr::summarise( 'mean(P)'=mean(Prediction),'sem(P)'=sd(Prediction)/sqrt(length(brks)) )

@@

rp = ggplot(endos.m, aes(x=brks, y=pred.r*100, ymin=(pred.r-`sem(P)`)*100, ymax=(pred.r+`sem(P)`)*100, group=brks)) + ylim(0,105) +
  geom_bar(stat='identity', fill='darkred', alpha=0.8, show.legend = F) + 
  geom_errorbar(color='grey21', width=0.2) +
  geom_text(aes(y=3, label=paste(round(pred.r*100), '%', sep='')), color='white') +
  geom_text(aes(y=pred.r*100-5, label=paste('n=',n.endo,'\ns=',n.samp,sep='')), color='white') + 
  labs(x='Months prior to endpoint', y='Percent predicted', title='Predictions prior to final endoscopy', subtitle='n = #endoscopies, s = #samples') +
  plot.theme + theme( axis.text.x=element_text(angle=45,hjust=1), text=element_text(size=12)) 
rp
ggsave(filename='plots/pred_endo.png', plot=rp, width=9, height=7, units='in', dpi=300)

#endos$months.before.final = endos$months.before.final*-1
#me = endos %>% group_by(Patient) %>% dplyr::summarise('max.endo' = max(i), 'months' = min(months.before.final))
```

`r pander(as.data.frame(pred.confidence))`

@@

@@

```{r, echo=F, warning=F, message=F, fig.width=5, fig.height=15 }
ga = grid.arrange(
  pred.tiles(subset(back, Patient %in% c(13,20,34,56,60,48) & Status == 'P'), ncol=1, cols=myPal, probs=T) + labs(title='Progressors') + theme(legend.position = 'none'),
  pred.tiles(subset(back, Patient %in% c(15,16,21,3,45,19) & Status == 'NP'), ncol=1, cols=myPal, probs=T) + labs(title='Non-Progressors') + theme(legend.position = 'none'),
  ncol=2)
ga

@@

p1 = pred.tiles(subset(back, Status == 'NP'), cols=myPal, probs=T, nrow=2) + labs(title='Non-Progressors') 
riskLegend = get.legend(p1)
p1 = p1 + theme(legend.position = 'bottom')
riskLegendLong = get.legend(p1)

p1 = p1 + theme(legend.position = 'none')
p2 = pred.tiles(subset(back, Status == 'P'), probs=T, cols=myPal,nrow=2) + labs(title='Progressors') + theme(legend.position = 'none')

ggsave('plots/pred_heatmap_ALL.png', plot=grid.arrange(p2,p1,nrow=2), width=52, height=16, units="in", limitsize=F, dpi = 300)
ggsave('plots/npg_heatmap.png', plot=p1, width=52, height=8, units="in", limitsize=F, dpi = 300)
ggsave('plots/pg_heatmap.png', plot=p2, width=52, height=8, units="in", limitsize=F, dpi = 300)
ggsave('plots/risk_legend.png', plot=grid.arrange(riskLegend), width=2,height=4, dpi=300)
ggsave('plots/risk_legend_long.png', plot=grid.arrange(riskLegendLong), width=8,height=2, dpi=300)
```

@@

```
One way of looking at the p53 prediction (e.g. Adam Bass) is to see how predictive the p-arm loss in chr17 is.
```{r, echo=F, warning=F, message=F, fig.height=5, fig.width=10}
ccgenes = read.table('~/Data/CosmicCensusGenes.tsv', sep='\t', header=T, stringsAsFactors=F)
ccgenes = ccgenes[-grep('^X|Y', ccgenes$Genome.Location),]
#ccgenes = ccgenes[-which(ccgenes$Mutation.Types == 'T'),] # drop the translocations
#ccgenes = subset(ccgenes, !grepl('^X|Y', Genome.Location))
#loc = unlist(strsplit(ccgenes$Genome.Location, ':|-'))
loc = do.call(rbind.data.frame, strsplit(ccgenes$Genome.Location, ':|-'))
colnames(loc) = c('chr','start','end')
loc[] = lapply(loc[], function(x) as.numeric(as.character(x)))

@@

>>>>

source('lib/load_patient_metadata.R')
source('lib/cv-pt-glm.R')
source('lib/common_plots.R')
source('lib/data_func.R')

@@

cache.dir = paste(data, 'Analysis/5e6_arms_all_exAHM0320', sep='/')

@@

sum.patient.data = subset(sum.patient.data, Hospital.Research.ID != 'AHM0320')

@@

ggplot(patient.info, aes(Status, Maximal, fill=Status)) + geom_boxplot() + labs(title='M') + plot.theme
ggplot(patient.info, aes(Status, Circumference, fill=Status)) + geom_boxplot() + labs(title='C') + plot.theme

@@

globalCN.plot = globalCN.plot + theme(legend.position = 'none') + labs(y='mean adj. seg.', title='All samples')
ggsave(filename='plots/globalCN.png', plot=globalCN.plot, width=16, height=4.6, units='in', dpi=300)
globalCN.plot

globalBECN.plot = globalBECN.plot + theme(legend.position = 'none') + labs(y='mean adj.seg.', title="Non-Dysplastic Barrett's Samples") 
ggsave(filename='plots/globalBECN.png', plot=globalBECN.plot, width=16, height=4.6, units='in', dpi=300)
globalBECN.plot

@@

tt = wilcox.test(subset(x, Status == 'Progressor')$cx, subset(x, Status == 'Non-Progressor')$cx)
pv = ggplot(x, aes(Status, cx, fill=Status)) + 
  geom_jitter(width=0.2, color='grey39') + 
  geom_boxplot(outlier.colour = NA, alpha=0.8) +
  labs(x='', y='cx', title='Per sample cx scores', subtitle=paste('p-value=', format.pval(tt$p.value, digits=2))) + theme(legend.position = 'none') +
  plot.theme
pv

@@

```{r noHGD, echo=F, message=F, warning=F, fig.width=6, fig.height=6}
rm(performance.at.1se, coefs, mp)

file = paste(cache.dir, 'nohgd.Rdata', sep='/')
if (file.exists(file)) {
  message(paste("loading file", file))
  load(file, verbose=T)
} else {
  stop(paste("No model file", file, "found"))
}
nohgd.coefs = coefs[[select.alpha]]
#do.call(grid.arrange, c(no.hgd.plots, top="No HGD/IMC samples", ncol=2))
ggsave('plots/nohgd_eyebrow.png', plot=do.call(grid.arrange, c(no.hgd.plots, top="No HGD/IMC samples", ncol=2)), height=10, width=10, units='in', dpi=300)

mp = model.performance(performance.at.1se, coefs)
mp$plot = mp$plot + labs(title='No HGD/IMC samples')
mp$plot

@@

```{r noLGD, echo=F, warning=F, message=F, fig.width=6, fig.height=6, eval=F}
file = paste(cache.dir, 'nolgd.Rdata', sep='/')
if (file.exists(file)) {
  load(file, verbose=T)
} else {
  stop(paste("No model file", file, "found"))
}

mp = model.performance(performance.at.1se, coefs)
mp$plot = mp$plot + labs(title='No LGD/HGD/IMC samples')
mp$plot

@@

```{r loonohgd, echo=F, message=F, warning=T, fig.width=5, fig.height=5, eval=F}
file = paste(cache.dir, 'loonohgd.Rdata', sep='/')
if (file.exists(file)) {
  message(paste("loading file", file))
  load(file, verbose=T)
} else {
  stop(paste("No file", file, "found"))
}

pg.sampNOHGD = do.call(rbind, pg.sampNOHGD)

@@

```{r, echo=F, fig.width=6, fig.height=6,eval=F}
ggplot(pg.sampNOHGD, aes(Prediction)) + geom_histogram(aes(fill=..x..), breaks=seq(0,1,0.1), show.legend = F) + scale_fill_distiller(palette = 'RdYlBu', name='P(P)') + labs(title='Prediction using model w/o HGD', y='n Samples', x='Probability') + plot.theme

@@

ggsave('plots/haz_coef.png', plot=p, height=5, width=5, units='in', dpi=300)

@@

```{r nohgd_coefs, message=F, warning=F, echo=F}
setdiff(ch$label, rownames(nohgd.coefs))
```

@@

ggsave('plots/global_cn_coefs.png', plot=p, width=16, height=4.6, units='in', dpi=300)

@@

ggsave('plots/ndbe_cn_coefs.png', plot=p, width=16, height=4.6, units='in', dpi=300)

@@

# oddPt = names(pg.samp)[which.max(performance.at.1se)]
# subset(sum.patient.data, Hospital.Research.ID == oddPt)
# pg.samp[[oddPt]]$Prediction
# 
# oddPt = names(pg.samp)[which.min(performance.at.1se)]

@@

load(file='~/Data/Ellie/Analysis/5e6_arms_all/insilico_splits.Rdata', verbose=T)

@@

multi.roc.plot(list(rocFirstEndo, rocLastEndo,rocEndoM, rocEndoA))

@@

multi.roc.plot(list(rocPtA, rocPtM))
```

@@

patient.info = subset(patient.info, Patient %in% sum.patient.data$Patient)

@@

#grid.arrange( tableGrob(annotation.table, rows=NULL, theme=tt3) )
pred.confidence$Risk = factor(pred.confidence$Risk, levels=c('Low','Moderate','High'), ordered = T)

lows = chisq.test(rbind(colSums(pred.confidence[1:5,c('NP','P')]),table(sum.patient.data$Status)))
highs = chisq.test(rbind(colSums(pred.confidence[6:10,c('NP','P')]),table(sum.patient.data$Status)))

mods = chisq.test(rbind(colSums(pred.confidence[4:7,c('NP','P')]),table(sum.patient.data$Status)))
```

```{r pcalib, echo=F, warning=F, message=F, fig.width=8, fig.height=6}
#pred.confidence$Risk[1:5] = 'Low'
#pred.confidence$Risk[6:10] = 'High'
pred.confidence$Risk[4:7] = 'Moderate'

pcal = ggplot(pred.confidence, aes(mn, perc)) + 
  geom_rect(aes(xmin=r1, xmax=r2, ymin=0,ymax=1, fill=Risk), alpha=0.6) + scale_fill_manual(values=riskCols, limits=levels(pred.confidence$Risk) ) +
  geom_point() + geom_abline(slope=1, intercept=0, color='grey39', linetype='dotted') + #geom_text(aes(label=paste(r1,r2,sep='-')), nudge_x=0.05) +
  geom_errorbar(aes(ymin=ci.low, ymax=ci.high), width=0.01) + labs(x='Mean prediction', y='Ratio of `P` patients', title='Prediction calibration') +
  scale_color_manual(values=riskCols,limits=levels(pred.confidence$Risk) ) + plot.theme + theme(legend.position = 'none')
pcal

@@

@@

unadjRR = ggplot(back, aes(OR)) + geom_histogram(aes(fill=..x..), bins=15, show.legend = F) +
  #scale_fill_distiller(palette = 'RdYlBu', type='div',name='P(P)') +
  scale_fill_gradientn(colors = myPal,  name='') + 
  labs(y='n Samples', x='Relative Risk', title='Unadjusted relative risk') + plot.theme
unadjRR
ggsave(filename='plots/unadj_relrisk.png', plot=unadjRR, width=6, height=5, units='in', dpi=300)

@@

ggsave(filename='plots/feature_jacknife.png', plot=p, width=6, height=8, unit='in', dpi=300)

@@

rp = ggplot(riskPath, aes(Pathology, ratio*100)) + 
  geom_bar(aes(group=Risk, fill=Risk),stat='identity', position=position_stack()) + 
  scale_fill_manual(values=riskCols, limits=levels(riskPath$Risk)) + 
  scale_color_manual(values=c('white','black','white'), limits=levels(pred.confidence$Risk) ) +
  geom_text(data=pathTotal, aes(label=paste('n=',nPath,sep=''), x=Pathology, y=101.5), position=position_stack(), size=4) +
  geom_text(data=subset(riskPath, ratio>0), aes(group=Risk,label=paste(ratio*100,'%',sep=''), color=Risk), position=position_stack(vjust = 0.5), size=5, show.legend = F) +
  facet_grid(~Status, scales='free_x', space = 'free_x') + plot.theme + 
  labs(title='Samples predicted by pathology', y='% Predicted', x='Pathology') + theme(legend.position = 'bottom')
rp

@@

ggplot(all.preds, aes(Patient, value/n.samples, group=variable, fill=variable)) +
  geom_bar(stat='identity', position = position_stack()) +
  scale_fill_manual(values=riskCols) +
  geom_text(aes(y=1.01, label=n.samples)) +
  facet_wrap(~Status, scales='free_x') + labs(y='Sample ratios', title='Per patient ratio of predictions') +
  plot.theme

@@

endos = back %>% group_by(Status, months.before.final, brks, PID) %>% dplyr::summarise(
  'pred.endo'=ifelse(length(which(Risk >= 'Moderate')) > 0, 1, 0), 
  'n.endos'=length(unique(PID)), 
  'n.samples'=length(Samplename),
  'mean.pred'=mean(Prediction),
  'pred.samples'=length(which(Risk >= 'Moderate')),
  'max.path'=max(Pathology),
  'Moderate'=length(which(Risk == 'Moderate')),
  'High'=length(which(Risk == 'High'))
  )

#endos$quant =  cut(endos$mean.pred, breaks = cuts)
#endos = merge(endos, pred.confidence[c('quant', 'Risk')], by='quant')

sq = c(0, 1, 1*12, 2*12, 4*12, 6*12, 8*12, 10*12, 15*12)
mlab = c('Endpoint','1-24','24-48','48','48-72','72-96','96-120','>120')

# months = 24
# sq = c(0, 1, 1*months, 2*months, 4*months, 6*months, 8*months)
# mlab = c()
# for (i in 1:length(sq)) {
#   mlab = c(mlab, paste(sq[i], sq[i+1], sep='-'))
# }
# mlab[1] = 'Endpoint'
# mlab = mlab[1:(length(sq)-1)]
# mlab[length(mlab)] = paste('>',sq[(length(sq)-1)],sep='')

endos$brks = cut(endos$months.before.final, include.lowest = T, ordered_result = T,breaks=sq,labels=mlab)
endos$brks = with(endos, factor(brks, rev(levels(brks)))) # invert order

endos.m = subset(endos, Status == 'P') %>% group_by(brks) %>% dplyr::summarise( 
  'pred.r'=sum(pred.endo)/sum(n.endos),'n.samp'=sum(n.samples), 'n.endo'=sum(n.endos), 
  'mod.r'=sum(Moderate)/sum(n.samples), 'high.r'=sum(High)/sum(n.samples))

tmp = subset(back, Status == 'P')
tmp$brks = cut(tmp$months.before.final, include.lowest = T, ordered_result = T,breaks=sq,labels=mlab)
tmp = tmp %>% group_by(brks) %>% dplyr::summarise( 'mean(P)'=mean(Prediction),'sem(P)'=sd(Prediction)/sqrt(length(brks)) )

@@

rp = ggplot(endos.m, aes(x=brks, y=pred.r*100, ymin=(pred.r-`sem(P)`)*100, ymax=(pred.r+`sem(P)`)*100, group=brks)) + ylim(0,105) +
  geom_bar(stat='identity', fill='orangered4', show.legend = F) + 
  geom_errorbar(color='grey21', width=0.2) +
  geom_text(aes(y=3, label=paste(round(pred.r*100), '%', sep='')), color='white') +
  geom_text(aes(y=pred.r*100-5, label=paste('n=',n.endo,'\ns=',n.samp,sep='')), color='white') + 
  labs(x='Months prior to endpoint', y='Percent predicted', title='Predictions prior to final endoscopy', subtitle='n = #endoscopies, s = #samples') +
  plot.theme + theme( axis.text.x=element_text(angle=45,hjust=1), text=element_text(size=12)) 
rp
ggsave(filename='plots/pred_endo.png', plot=rp, width=9, height=7, units='in', dpi=300)

#endos$months.before.final = endos$months.before.final*-1
#me = endos %>% group_by(Patient) %>% dplyr::summarise('max.endo' = max(i), 'months' = min(months.before.final))
```
# TODO number endoscopies so the images are scaled 
`r pander(as.data.frame(pred.confidence))`

@@

table(rev(arrange(subset(back, Patient == 19), brks)$brks))

@@

```{r, echo=F, warning=F, message=F, fig.width=5, fig.height=15 }
#back = back %>% group_by(Patient) %>% dplyr::mutate(endo.num = as.integer(droplevels(brks)) )
#back$endo.num = factor(back$endo.num, ordered = T)

#back$endo.num = factor(as.integer(back$brks))
#back = back %>% group_by(Patient) %>% dplyr::mutate(endo.num = ifelse(endo.num == max(endo.num), 'End', endo.num)   )

ga = grid.arrange(
  pred.tiles(subset(back, Patient %in% c(13,20,34,56,60,48) & Status == 'P'), ncol=1, cols=myPal, probs=T) + labs(title='Progressors') + theme(legend.position = 'none'),
  pred.tiles(subset(back, Patient %in% c(15,16,21,3,45,19) & Status == 'NP'), ncol=1, cols=myPal, probs=T) + labs(title='Non-Progressors') + theme(legend.position = 'none'),
  ncol=2)
ga

@@

p1 = pred.tiles(subset(back, Status == 'NP'), cols=myPal, probs=T, nrow=2, scales='free', drop=T) + labs(title='Non-Progressors') 
#p1 = pred.tiles(subset(back, Status == 'NP'), cols=riskCols, probs=F, OR=F, nrow=2) + labs(title='Non-Progressors') 
riskLegend = get.legend(p1)
p1 = p1 + theme(legend.position = 'bottom')
riskLegendLong = get.legend(p1)

p1 = p1 + theme(legend.position = 'none')
p2 = pred.tiles(subset(back, Status == 'P'), probs=T, cols=myPal,nrow=2,scales='free',drop=T) + labs(title='Progressors') + theme(legend.position = 'none')

ggsave('plots/pred_heatmap_ALL.png', plot=grid.arrange(p2 + labs(x='', y='',title=''),p1 + labs(x='', y='',title=''),nrow=2), width=11, height=3.4, units="in", limitsize=F, dpi = 300, scale = 3)
ggsave('plots/npg_heatmap.png', plot=p1, width=52, height=8, units="in", limitsize=F, dpi = 300)
ggsave('plots/pg_heatmap.png', plot=p2, width=52, height=8, units="in", limitsize=F, dpi = 300)
ggsave('plots/risk_legend.png', plot=grid.arrange(riskLegend), width=2,height=4, dpi=300)
ggsave('plots/risk_legend_long.png', plot=grid.arrange(riskLegendLong), width=8,height=2, dpi=300)

p1 = pred.tiles(subset(back, Status == 'NP'), cols=myPal, probs=T, nrow=2, scales='free_x', drop=T) + labs(title='Non-Progressors') + theme(legend.position = 'none') 
p2 = pred.tiles(subset(back, Status == 'P'), probs=T, cols=myPal,nrow=2,scales='free_x',drop=T) + labs(title='Progressors') + theme(legend.position = 'none')
ggsave('plots/pred_heatmap_ALL_widthadj.png', plot=grid.arrange(p2 + labs(x='', y='',title=''),p1 + labs(x='', y='',title=''),nrow=2), width=11, height=3.4, units="in", limitsize=F, dpi = 300, scale = 3)
```

@@

One way of looking at the p53 prediction (e.g. Adam Bass) is to see how predictive the p-arm loss in chr17 is.
```{r, echo=F, warning=F, message=F, fig.height=5, fig.width=10}
ccgenes = read.table('~/Data/CosmicCensusGenes.tsv', sep='\t', header=T, stringsAsFactors=F)
ccgenes = ccgenes[-grep('^X|Y', ccgenes$Genome.Location),]
#ccgenes = ccgenes[-which(ccgenes$Mutation.Types == 'T'),] # drop the translocations
#ccgenes = subset(ccgenes, !grepl('^X|Y', Genome.Location))
#loc = unlist(strsplit(ccgenes$Genome.Location, ':|-'))
loc = do.call(rbind.data.frame, strsplit(ccgenes$Genome.Location, ':|-'))
colnames(loc) = c('chr','start','end')
loc[] = lapply(loc[], function(x) as.numeric(as.character(x)))

@@

# Validation

## Downsampled Datasets

```{r echo=F, eval=T}
dirA = '~/Data/Ellie/Analysis/downsampled5G_BEAdjacent'
dirB = '~/Data/Ellie/Analysis/downsampled5G_NDBE'

file = paste(cache.dir, 'model_data.Rdata', sep='/')
if (!file.exists(file))
  stop(paste("Missing data file", file))
load(file, verbose=T)
file = paste(cache.dir, 'all.pt.alpha.Rdata', sep='/')
if (!file.exists(file))
  stop(paste("Missing data file", file))
load(file, verbose=T)
rm(dysplasia.df,coefs,plots,labels)

select.alpha = '0.9'
fitV = models[[select.alpha]]
lambda.opt = performance.at.1se[[select.alpha]][, 'lambda']

predA = list.files(dirA, 'predictions_all_exAHM0320.Rdata', full.names = T)
predB = list.files(dirB, 'predictions_all_exAHM0320.Rdata', full.names = T)

files = c()
if (length(predA) < 1)
  files = c(list.files(dirA, pattern='raw', full.names=T))
if (length(predB) < 1)
  files = c(files, list.files(dirB, pattern='raw', full.names=T))

if (length(files) > 0) {
  sampleNames = sub('_raw.txt', '', basename(files))
  dspred = data.frame(matrix(ncol=2, nrow=length(sampleNames), dimnames=list(sampleNames,c('Prob','RR'))))

  predA = sub('_raw.txt','',basename(c(list.files(dirA, pattern='raw', full.names=T))))
  predB = sub('_raw.txt','',basename(c(list.files(dirB, pattern='raw', full.names=T))))

  for (n in 1:length(files)) {
    print(n)
    print(sampleNames[n])

    rs = read.table(files[n], sep='\t',header=T)
    if (ncol(rs) < 4)
      stop("File needs to contain 4 columns: chr, start, end, sample_value")

    segs <- tile.segmented.data(rs, size=5e6, chr.info=chr.lengths)
    segM = as.matrix(segs[,-c(1:3)])
    rownames(segM) = paste(segs$chr, ':', segs$start, '-', segs$end, sep='')
    segM = t(segM)
    for (i in 1:ncol(segM)) 
      segM[,i] = unit.var(segM[,i], z.mean[i], z.sd[i])

    arms <- tile.segmented.data(rs, size='arms', chr.info=chr.lengths)
    armsM = as.matrix(arms[,-c(1:3)])
    rownames(armsM) = paste(arms$chr, ':', arms$start, '-', arms$end, sep='')
    armsM = t(armsM)
    for (i in 1:ncol(armsM)) 
      armsM[,i] = unit.var(armsM[,i], z.arms.mean[i], z.arms.sd[i])

    nrow(armsM) == nrow(segM)

    cx.score = score.cx(segM,1)

    mergedDf = subtract.arms(segM, armsM)
    mergedDf = cbind(mergedDf, 'cx' = unit.var(cx.score, mn.cx, sd.cx))
    dim(mergedDf)

    # Predict function giving me difficulty when I have only a single sample, this ensures the dimensions are the same
    sparsed_test_data <- Matrix(data=0, nrow=nrow(mergedDf),  ncol=ncol(mergedDf),
                                dimnames=list(rownames(mergedDf),colnames(mergedDf)), sparse=T)
    for(col in colnames(mergedDf)) sparsed_test_data[,col] = mergedDf[,col]

    preds = predict(fitV, newx=sparsed_test_data, s=lambda.opt, type='response')
    RR = predict(fitV, newx=sparsed_test_data, s=lambda.opt, type='link')

    dspred[n,] = c(preds[,1],RR[,1])
    print(dspred[1:n,])
  }
    dspred$Type = NA
    dspred[predA,'Type'] = 'AdjacentBE'
    dspred[predB,'Type'] = 'BEST2 NDBO'
    pred = dspred

    dspred = subset(pred, Type == 'AdjacentBE')

    save(dspred, file=paste(dirA, 'predictions_all_exAHM0320.Rdata', sep='/'))

    dspred = subset(pred, Type == 'BEST2 NDBO')
    save(dspred, file=paste(dirB, 'predictions_all_exAHM0320.Rdata', sep='/'))

    dspred = pred

} else {
  load(predB, verbose=T)
  dspredB = dspred

  load(predA, verbose=T)

  dspred = rbind(dspred, dspredB)
  dspred$Type = NA

  dspred[sub('_raw.*','',basename(list.files(dirA, pattern='raw', full.names=T))),'Type'] = 'AdjacentBE'
  dspred[sub('_raw.*','',basename(list.files(dirB, pattern='raw', full.names=T))),'Type'] = 'BEST2 NDBO'

}

ggplot(dspred, aes(Type, RR, fill=Type)) + geom_boxplot() + geom_jitter(width=0.2) +
  labs(x='', y='Relative Risk', title='Predictions on downsampled patients') + 
  theme(legend.position = 'none') + plot.theme

adj = subset(dspred, Type == 'AdjacentBE')
prg = back[,c('Prediction', 'OR', 'Status')]
prg$Status = ifelse(prg$Status == 'P', 'Progressor','Non-Progressor')

#  subset(back, Status == 'P', select=c('Prediction', 'OR', 'Status'))
colnames(prg) = colnames(adj)
m = rbind(adj,prg)
m$Type = factor(m$Type, levels=c('Non-Progressor','BEST2 NDBO','Progressor','AdjacentBE'), ordered = T)

pr = ggplot(m, aes(Type, RR, fill=Type)) + 
  geom_jitter(width=0.2) + geom_boxplot() + 
  labs(x='', y='RR', title='Predicted Relative Risk') + 
  theme(legend.position = 'none', text=element_text(size=11)) + plot.theme 
pr

ggsave('plots/adjacentBE.png', plot=pr, width=8, height=8,dpi=300)

# Clinical Story

First note here is that our probability distribution for these 20 patients is somewhat flatter than in the training set.  Likely a consequence of sample size. The labels show the proprotions of progressors that fell into each quintile in the training set.

## Vignettes

```{r, echo=F, warning=F, message=F, fig.width=5, fig.height=5}
vign.pts = as_tibble(subset(sum.patient.data, Patient %in% c(3, 33, 34, 49, 56, 20), select=c('Status','Patient','age.diagnosis', 'smoking', 'C','M', 'initial.endo','final.endo', 'total.endos','total.samples')))
colnames(vign.pts) = c('Status','Patient','Age at Diagnosis','Smoker','Circumference', 'Maximal length', 'Intial Endoscopy','Final Endoscopy', 'Total Endoscopies','Total Samples')

## P
p = vig.plot(df=subset(back, Patient == 34),info= subset(patient.info, Patient == 34),cols=riskCols) + theme(legend.position = 'bottom')
vigLeg = get.legend(p)
p = p + theme(legend.position = 'none')
ggsave('plots/vignette/vig_34.png', plot=p, width=6, height=6,dpi=300)

p = vig.plot(subset(back, Patient == 56), subset(patient.info, Patient == 56),cols=riskCols) + theme(legend.position = 'none')
ggsave('plots/vignette/vig_56.png', plot=p, width=6, height=6,dpi=300)

p = vig.plot(subset(back, Patient == 20), subset(patient.info, Patient == 20),cols=riskCols) + theme(legend.position = 'none')
ggsave('plots/vignette/vig_20.png', plot=p, width=6, height=6,dpi=300)
